---
title: "Data analysis with clustRomics"
author: "Fulvia D'Ercole"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data analysis with clustRomics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

#Introduction


```{r, message=FALSE, warning=FALSE}
library(clustRomics)
library(SummarizedExperiment)
```

# 1. Loading data 
# To use the example file included in the package:

``` {r}
example_path <- system.file("extdata", "test_genes.csv", package = "clustRomics")
se <- read_expression(example_path)
``` 

# Or load your own data:

``` {r, eval = FALSE}
se <- read_expression("path/to/your/data.csv")
```

# 2. Normalization: You can choose between different normalization strategies depending on your downstream analysis needs

```{r}
#Methods available: "log" (log-transformation) or "quantile"
se <- normalize_data(se, method = "log")
```

# 3. Genes cluster: group genes with similar expression profiles using various algorithms:

```{r}
#Methods available: "kmeans", "hclust" (hierarchical)
se <- cluster_genes(se, method = "kmeans", k = 5)
```

# 4. Analysis and visualization
## Detecting Outlier: Identify and remove genes with anomalous distributions:

```{r}
se <- detect_outliers(se, threshold = 2)
```

# Heatmap
## View gene clusters across different conditions:

``` {r}
plot_expression_heatmap(se)
```

# PCA Plot 
## Explore the distribution of samples in principal component space:

```{r}
plot_pca(se)
```

# Biological Enrichment (GSEA) 
## Identify biological pathways (GOs) enriched in a specific cluster:

```{r}
enrich_res <- run_cluster_enrichment(se, cluster_id = 1)
```